---
title: "Time_Plots_and_ASTSA"
output: html_document
---

Some time series from the ASTSA package:
- jj, flu, gtemp_both, gtemp_land, star

```{r}
library(astsa)
## require(astsa) does basically the exact same thing
help(astsa)
```
Now looking at JJ quarterly earnings per share
```{r}
help(jj)
```
Now plotting
```{r}
plot(jj, type='o', main='J&J Quarterly Earnings per Share', ylab='Earnings', xlab='Years')
```

Pneumonia and influenze deaths
```{r}
plot(flu, main='Monthly Pneumonia and Influenze Deaths in US', ylab='Number of Deaths per 10,000 people', xlab='Months')
```
There's probably some seasonality going on here, but there might also be an overall trend of deaths going down.

Land-ocean temperature deviations (update is called gtemp_both (original was called globtemp (and globtempl for land)))
```{r}
plot(gtemp_both, main='Global mean land-ocean deviations from average temp from 1880-2010', xlim=c(1880, 2010), ylab='Temp deviations', type='o', xlab='Years')
```

Now for land only
```{r}
plot(gtemp_land, main='Global mean land deviations from average temp from 1880-2010', xlim=c(1880, 2010), ylab='Temp deviations', type='o', xlab='Years')
```

Variable Star (Magnitude of star taken at midnight)
```{r}
plot(star, main='Magnitude of star taken at midnight for 600 consec. days', ylab='Magnitude', xlab='Days')
```



(Weak) Stationarity
Things we want:
- No systematic change in mean (i.e., no trend)
- No systematic change in variance
- No periodic fluctuations

Basically, we want the properties of one section of the data to be like the properties of the other sections of the data.

For a non-stationary time series (which is usually the case), we'll do some transformations to get stationary time series.















