---
title: "Basic_Stats_Review"
output: html_document
---

Review of Basic Statistics 1 - Simple Linear Regression
```{r}
plot(co2, main='Atmosphereic CO2 Concentration')
```
```{r}
# Developing a linear model
co2.linear.model = lm(co2 ~ time(co2))
co2.linear.model
```
```{r}
plot(co2, main='Atmospheric CO2 Concentration with Fitted Line')
abline(co2.linear.model)
```
Quiz
```{r}
x=c(1,2,3,4);
y=c(5, 7, 12, 13);
m=lm(y~x)
m
```

Review of Basic Statistics 2 - More Linear Regression
```{r}
co2.lm = lm(co2 ~ time(co2))
co2.lm
```
```{r}
## plot(co2, main='Atmospheric CO2 Concentration with Fitted Line')
## abline(co2.lm)

# Getting residuals
co2.residuals = resid(co2.lm)
hist(co2.residuals, main="Histogram of Residuals")
```
This looks relatively normal, but let's use qq plots instead
```{r}
qqnorm(co2.residuals)
qqline(co2.residuals)
```
So now we see that there's systematic departures from normality, so there might be some cause for concern here.
```{r}
co2_time_plot = co2.residuals ~ time(co2)
plot(co2_time_plot)
```
There seems to be some systematic changes in residuals. Let's zoom in to see if we can see anything further.
```{r}
plot(co2_time_plot, xlim=c(1960, 1963), main="Zoomed Residuals")
```

Quiz
```{r}
library(faraway)
plot(punting$Distance~ punting$Hang); 
abline( lm(punting$Distance~ punting$Hang) )

m=lm(punting$Distance~ punting$Hang)
qqnorm(resid(m))
qqline(resid(m))
# Overall, it looks at least relatively normal
```

Review of Basic Statistics 3 - Inference
```{r}
# Gossett Data (10 people)
help(sleep)
```

```{r}
plot(extra ~ group, data=sleep, main='Extra Sleep in Gossett Data by Group')
attach(sleep)  # Just makes the variables available in the environment
extra.1=extra[group==1]
extra.2=extra[group==2]
```
T-tests
```{r}
t.test(extra.1, extra.2, paired=TRUE, alternative="two.sided")
```
Because we got a t-statistic of -4.06 and a CI that doesn't include 0, we can pretty confidently rejected the null hypothesis.

Review of Basic Statistics 4 - Correlation
```{r}
# Using the trees example
help(trees)
```
Using a pairplot to see correlation
```{r}
pairs(trees, pch=21, bg=c('red'))  # pch and bg are just color things
```
Clearly, girth is a good predictor of volume.
```{r}
# Calculating the covariance
cov(trees)
```
This doesn't look great because of the units involved, so we'll look at the correlation to account for that.
```{r}
cor(trees)
```
Correlation can be expressed rather simply as:
r = (SSXY) / ( sqrt(SSX) * sqrt(SSY) )



















